### 2020-01-29 wuhan.tex

Since December 2019, the novel coronavirus (2019-nCoV) has been
spreading in China and in other parts of the world.
As of January 28th, 2020, the World Health Organization has
confirmed 4593 cases, including 56 confirmed cases in 14 different
countries, outside China \citep{who28report}.
Although the virus is believed to have originated
from animal reservoirs, a recent case from Viet Nam
demonstrated its ability to directly transmit between
humans \citep{who26report},
posing a greater threat for its spread.

As the disease continues to spread, 
many disease modelers have been rushing to publish their 
analysis of the outbreak and, in particular, their
estimates of the basic reproductive number $\mathcal R_0$ (i.e., the 
average number of secondary cases generated 
by a primary case in a fully susceptible population \citep{anderson1991infectious, diekmann1990definition}).
The basic reproductive number is of particular interest 
because it allows prediction about the final size of an epidemic \citep{anderson1991infectious, ma2006generality, arino2007final, andreasen2011final, miller2012note}.
While their efforts are valuable, it is difficult to
compare their estimates of $\mathcal R_0$ because their methods 
and assumptions, which can directly affect their estimates and
the associated uncertainties, vary widely.

Here, we present a simple, statistical framework for evaluating uncertainties 
associated with the estimates of $\mathcal R_0$
across a wide range of models and compare estimates of $\mathcal R_0$ for the outbreak.
Our results indicate that most early estimates of $\mathcal R_0$ for the outbreak
are likely to be overly confident.

>>>>

Since December 2019, the novel coronavirus (2019-nCoV) has been
spreading in China and other parts of the world.
As of January 28th, 2020, the World Health Organization has
confirmed 4593 cases, including 56 confirmed cases in 14 different
countries, outside China \citep{who28report}.
Although the virus is believed to have originated
from animal reservoirs, a recent case from Viet Nam
demonstrated its ability to directly transmit between
humans \citep{who26report},
posing a greater threat for its spread.

As the disease continues to spread, 
many researchers have been rushing to publish their 
analysis of the outbreak and, in particular, their
estimates of the basic reproductive number $\mathcal R_0$ (i.e., the 
average number of secondary cases generated 
by a primary case in a fully susceptible population \citep{anderson1991infectious, diekmann1990definition}).
Estimating the basic reproductive number is of particular interest during an outbreak
because it allows prediction about the final size of the outbreak \citep{anderson1991infectious, ma2006generality, arino2007final, andreasen2011final, miller2012note} and the required level of intervention \citep{anderson1991infectious}.
However, it can be difficult to assess the validity of the estimates of $\mathcal R_0$ (as well as the associated degrees of uncertainty)
when the estimation methods and their underlying assumptions vary widely.

Here, we present a simple, statistical framework for evaluating uncertainties 
associated with the estimates of $\mathcal R_0$
across a wide range of models using a generation-interval based approach and compare early estimates of $\mathcal R_0$ of the 2019-nCoV outbreak.
Our results indicate that most early estimates of $\mathcal R_0$ for the outbreak
are likely to be overly confident.
